---
title: "ENCLOSE: Data Harmonization"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
    toc_depth: 3
    code_folding: show
  pdf_document: default
---
```{r packages}
# Packages
if (!requireNamespace('dplyr', quietly = TRUE)) install.packages('dplyr')
library(dplyr)
if (!requireNamespace('haven', quietly = TRUE)) install.packages('haven')
library(haven)
if (!requireNamespace('labelled', quietly = TRUE)) install.packages('labelled')
library(labelled)
if (!requireNamespace('mirt', quietly = TRUE)) install.packages('mirt')
library(mirt)
if (!requireNamespace('stringr', quietly = TRUE)) install.packages('stringr')
library(stringr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```


```{r packages-001, include=TRUE, message=FALSE, warning=FALSE}
library(haven)
library(dplyr)
library(labelled)
library(stringr)
library(mirt)
```

```{r packages-002}
ESS10 <- read_sav("ESS10.sav")
ZA7649 <- read_sav("ZA7649_v2-1-0.sav")

# initialize datasets
ESS10_h=ESS10      # _h stand for harmonized
ZA7649_h=ZA7649    # _h stand for harmonized
```

# COMMON VARIABLES

## OBSERVED COMMON VARIABLES
### Age
```{r age-003}
#summary(ESS10$agea)
ESS10_h <- ESS10_h[!is.na(ESS10_h$agea), ]
#summary(ZA7649$d11)
ZA7649_h <- ZA7649_h[!is.na(ZA7649_h$d11), ]
```

### Gender
```{r gender-004}
ESS10_h=subset(ESS10_h, gndr %in% c(1,2))
ESS10_h$gndr <- labelled(ESS10_h$gndr,
                         c("Man" = 1,
                           "Woman" = 2
                         ))

# split missing values of ZA7649$d10
ZA7649_h=subset(ZA7649_h, d10 %in% c(1,2))

# Recoding values and labels of ZA7649_h$d10
ZA7649_h$d10 <- labelled(ZA7649_h$d10,
                            c("Man" = 1,
                              "Woman" = 2))
```

### Occupation
```{r occupation-005}
# split missing values of ESS10$mnactic
ESS10_h=subset(ESS10_h, mnactic %in% c(1,2,3,4,5,6,7,8))

# Recoding values and labels of ESS10$mnactic
ESS10_h=ESS10_h %>%
mutate(mnactic = case_when(
    mnactic %in% c(1,7) ~ 1,
    mnactic %in% c(3,4,8) ~ 2,
    mnactic %in% c(5,6) ~ 3,
    mnactic == 2 ~ 4
  ))
ESS10_h$mnactic <- labelled(ESS10_h$mnactic,
        c("Employed" = 1,
          "Unemployed" = 2,
          "Retired" = 3,
          "Education" = 4))

# split missing values of ZA7649$d15a
ZA7649_h=subset(ZA7649_h, d15a %in% c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18))

# Recoding values and labels of ZA7649$d15a
ZA7649_h=ZA7649_h %>%
  mutate(d15a = case_when(
    d15a %in% c(1,5,6,7,8,9,10,11,12,13,14,15,16,17,18) ~ 1,
    d15a == 3 ~ 2,
    d15a == 4 ~ 3,
    d15a == 2 ~ 4
  ))
ZA7649_h$d15a <- labelled(ZA7649_h$d15a,
                            c("Employed" = 1,
                              "Unemployed" = 2,
                              "Retired" = 3,
                              "Education" = 4))

```


### Domicile
```{r domicile-006}
# split missing values of ESS10$domicil
ESS10_h=subset(ESS10_h, domicil %in% c(1,2,3,4,5))

# Recoding values and labels of ESS10$domicil
ESS10_h=ESS10_h %>%
  mutate(domicil = case_when(
    domicil %in% c(1,2) ~ 1,
    domicil == 3 ~ 2,
    domicil %in% c(4,5) ~ 3
  ))
ESS10_h$domicil <- labelled(ESS10_h$domicil,
                             c("A big city or large town" = 1,
                               "Small middle town" = 2,
                               "Rural Area or village" = 3))
# split missing values of ZA7649$d25
ZA7649_h=subset(ZA7649_h, d25 %in% c(1,2,3))

# Recoding values and labels of ZA7649$d25
ZA7649_h=ZA7649_h %>%
  mutate(d25 = case_when(
    d25 == 3 ~ 1,
    d25 == 2 ~ 2,
    d25 == 1 ~ 3
  ))
ZA7649_h$d25 <- labelled(ZA7649_h$d25,
                         c("A big city or large town" = 1,
                           "Small middle town" = 2,
                           "Rural Area or village" = 3))

```

### Subjective income: economic difficulties
```{r subjective-income-economic-difficulties-007}
# split missing values of ESS10$hincfel
ESS10_h=subset(ESS10_h, hincfel %in% c(1,2,3,4))

# Recoding values and labels of ESS10$hincfel
ESS10_h=ESS10_h %>%
  mutate(hincfel = case_when(
    hincfel %in% c(1,2) ~ 1,
    hincfel %in% c(3,4) ~ 2
  ))
ESS10_h$hincfel <- labelled(ESS10_h$hincfel,
                            c("No" = 1,
                              "Yes" = 2))

# split missing values of ZA7649$d60
ZA7649_h=subset(ZA7649_h, d60 %in% c(1,2,3))

# Recoding values and labels of ZA7649$d60
ZA7649_h=ZA7649_h %>%
  mutate(d60 = case_when(
    d60 == 3 ~ 1,
    d60 %in% c(1,2) ~ 2
  ))
ZA7649_h$d60 <- labelled(ZA7649_h$d60,
                         c("No" = 1,
                              "Yes" = 2))

```


### Life Satisfaction
```{r life-satisfaction-008}
# split missing values of ESS10$stflife
ESS10_h=subset(ESS10_h, stflife %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$stflife
ESS10_h=ESS10_h %>%
  mutate(stflife = case_when(
    stflife %in% c(0,1,2) ~ 1,
    stflife %in% c(3,4,5) ~ 2,
    stflife %in% c(6,7,8) ~ 3,
    stflife %in% c(9,10) ~ 4
  ))
ESS10_h$stflife <- labelled(ESS10_h$stflife,
                            c("Not at all satisfied" = 1,
                              "Not very satisfied" = 2,
                              "Fairly satisfied" = 3,
                              "Very satisfied" = 4))


# split missing values of ZA7649$d70
ZA7649_h=subset(ZA7649_h, d70 %in% c(1,2,3,4))

# Recoding values and labels of ZA7649$d70
ZA7649_h=ZA7649_h %>%
  mutate(d70 = case_when(
    d70  == 4 ~ 1,
    d70  == 3 ~ 2,
    d70  == 2 ~ 3,
    d70  == 1 ~ 4
  ))
ZA7649_h$d70 <- labelled(ZA7649_h$d70,
                         c("Not at all satisfied" = 1,
                           "Not very satisfied" = 2,
                           "Fairly satisfied" = 3,
                             "Very satisfied" = 4))

```


### Satisfaction with national economy
```{r satisfaction-with-national-economy-009}
# split missing values of ESS10$stfeco
ESS10_h=subset(ESS10_h, stfeco %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$stfeco
ESS10_h=ESS10_h %>%
  mutate(stfeco = case_when(
    stfeco %in% c(0,1,2) ~ 1,
    stfeco %in% c(3,4,5) ~ 2,
    stfeco %in% c(6,7,8) ~ 3,
    stfeco %in% c(9,10) ~ 4
  ))
ESS10_h$stfeco <- labelled(ESS10_h$stfeco,
                           c("Very bad" = 1,
                             "Rather bad" = 2,
                             "Rather good" = 3,
                             "Very good" = 4))

# split missing values of ZA7649$qa1a_2
ZA7649_h=subset(ZA7649_h, qa1a_2 %in% c(1,2,3,4))

# Recoding values and labels of ZA7649$qa1a_2
ZA7649_h=ZA7649_h %>%
  mutate(qa1a_2 = case_when(
    qa1a_2 ==4 ~ 1,
    qa1a_2 ==3 ~ 2,
    qa1a_2 ==2 ~ 3,
    qa1a_2 ==1 ~ 4
  ))
ZA7649_h$qa1a_2 <- labelled(ZA7649_h$qa1a_2,
                             c("Very bad" = 1,
                               "Rather bad" = 2,
                               "Rather good" = 3,
                               "Very good" = 4))
```

### Attachment to country
```{r attachment-to-country-010}
# split missing values of ESS10$atchctr
ESS10_h=subset(ESS10_h, atchctr %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$atchctr
ESS10_h=ESS10_h %>%
  mutate(atchctr = case_when(
    atchctr %in% c(0,1,2) ~ 1,
    atchctr %in% c(3,4,5) ~ 2,
    atchctr %in% c(6,7,8) ~ 3,
    atchctr %in% c(9,10) ~ 4
  ))
ESS10_h$atchctr <- labelled(ESS10_h$atchctr,
                           c("Not at all attached" = 1,
                             "Not very attached" = 2,
                             "Fairly attached" = 3,
                             "Very attached" = 4))


# split missing values of ZA7649$qc1a_2
ZA7649_h=subset(ZA7649_h, qc1a_2 %in% c(1,2,3,4))

# Recoding values and labels of ZA7649$qa1a_2
ZA7649_h=ZA7649_h %>%
  mutate(qc1a_2 = case_when(
    qc1a_2 ==4 ~ 1,
    qc1a_2 ==3 ~ 2,
    qc1a_2 ==2 ~ 3,
    qc1a_2 ==1 ~ 4
  ))
ZA7649_h$qc1a_2 <- labelled(ZA7649_h$qc1a_2,
                            c("Not at all attached" = 1,
                              "Not very attached" = 2,
                              "Fairly attached" = 3,
                              "Very attached" = 4)
                            )
```


### Political position
```{r political-position-011}
# split missing values of ESS10$lrscale
ESS10_h=subset(ESS10_h, lrscale %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$lrscale
ESS10_h=ESS10_h %>%
  mutate(lrscale = case_when(
    lrscale %in% c(0,1,2,3) ~ 1,
    lrscale  %in% c(4,5,6,7) ~ 2,
    lrscale  %in% c(8,9,10) ~ 3
  ))
ESS10_h$lrscale <- labelled(ESS10_h$lrscale,
                            c("Left" = 1,
                              "Centre" = 2,
                              "Right" = 3))


# split missing values of ZA7649$d1
ZA7649_h=subset(ZA7649_h, d1 %in% c(1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ZA7649$d1
ZA7649_h=ZA7649_h %>%
  mutate(d1 = case_when(
    d1 %in% c(1,2,3) ~ 1,
    d1  %in% c(4,5,6,7) ~ 2,
    d1  %in% c(8,9,10) ~ 3
  ))
ZA7649_h$d1 <- labelled(ZA7649_h$d1,
                        c("Left" = 1,
                          "Centre" = 2,
                          "Right" = 3))
```

## Citizenship
```{r citizenship-012}
# split missing values of ESS10$ctzcntr
ESS10_h=subset(ESS10_h, ctzcntr %in% c(1,2))

# Recoding values and labels of ESS10$ctzcntr
ESS10_h=ESS10_h %>%
  mutate(ctzcntr = case_when(
    ctzcntr == 1 ~ 1,
    ctzcntr == 2 ~ 0
  ))
ESS10_h$ctzcntr <- labelled(ESS10_h$ctzcntr,
                            c("No" = 0,
                              "Yes" = 1))

# 1. Identifica le colonne che iniziano con "q1."
colonne_q1_names <- names(ZA7649_h)[startsWith(names(ZA7649_h), "q1.")]

# 2. Crea la tabella di corrispondenza tra nomi delle nazionalità e codici ISO alpha-2
corrispondenze_iso <- tibble(
  nazionalita = c("BELGIUM", "DENMARK", "GERMANY", "GREECE", "SPAIN", "FRANCE",
                  "IRELAND", "ITALY", "LUXEMBOURG", "NETHERLANDS", "PORTUGAL",
                  "UNITED KINGDOM", "AUSTRIA", "SWEDEN", "FINLAND", "CYPRUS",
                  "CZECH REPUBLIC", "ESTONIA", "HUNGARY", "LATVIA", "LITHUANIA",
                  "MALTA", "POLAND", "SLOVAKIA", "SLOVENIA", "BULGARIA", "ROMANIA",
                  "CROATIA", "TURKEY", "CYPRUS - TCC", "NORTH MACEDONIA",
                  "MONTENEGRO", "SERBIA", "ALBANIA", "OTHER COUNTRIES", "DK"),
  iso_alpha2 = c("BE", "DK", "DE", "GR", "ES", "FR",
                 "IE", "IT", "LU", "NL", "PT",
                 "GB", "AT", "SE", "FI", "CY",
                 "CZ", "EE", "HU", "LV", "LT",
                 "MT", "PL", "SK", "SI", "BG", "RO",
                 "HR", "TR", "TR", "MK",
                 "ME", "RS", "AL", NA, NA),
  colonna_q1 = paste0("q1.", 1:36)
)

# 3. Itera attraverso le colonne q1 e crea le colonne ISO
for (i in seq_along(colonne_q1_names)) {
  col_q1_name <- colonne_q1_names[i]
  etichetta <- names(ZA7649_h)[which(names(ZA7649_h) == col_q1_name)]

  codice_iso <- corrispondenze_iso %>%
    filter(colonna_q1 == etichetta) %>%
    pull(iso_alpha2)

  nuova_colonna_nome <- paste0("iso_", col_q1_name)

  ZA7649_h <- ZA7649_h %>%
    mutate(!!nuova_colonna_nome := ifelse(as.numeric(!!sym(col_q1_name)) == 1, codice_iso, NA_character_))
}

# 4. Identifica le colonne iso_q1.x
iso_colonne <- names(ZA7649_h)[startsWith(names(ZA7649_h), "iso_q1.")]

# 5. Crea la colonna ctzcntr
ZA7649_h <- ZA7649_h %>%
  mutate(
    ctzcntr = case_when(
      is.na(isocntry) ~ 0,
      TRUE ~ as.integer(rowSums(select(., all_of(iso_colonne)) == isocntry, na.rm = TRUE) > 0)
    )
  )

# 6. Elimina le colonne iso_q1.x e inserisci le labels
ZA7649_h <- ZA7649_h %>%
  select(-all_of(iso_colonne))%>%
  set_variable_labels(
    ctzcntr = "Country"
  )

ZA7649_h$ctzcntr <- labelled(ZA7649_h$ctzcntr,
                            c("No" = 0,
                              "Yes" = 1))



#harmonizing countries ZA7649_h$isocntry with ESS10_h$cntry
ZA7649_h$iso_h=ZA7649_h$isocntry
ZA7649_h$iso_h[ZA7649_h$iso_h %in% c("DE-W" , "DE-E"  )]  = "DE"
ZA7649_h$iso_h[ZA7649_h$iso_h %in% c("CY-TCC"  )] = 'CY'
ZA7649_h$iso_h[ZA7649_h$iso_h %in% c("GB-NIR"  , "GB-UKM"  )] = "GB"


# Clean Workspace
rm(codice_iso, col_q1_name, colonne_q1_names, corrispondenze_iso, etichetta, i, iso_colonne, nuova_colonna_nome)

```


#### Selection of citizens given the small percentage of non citizens
```{r selection-of-citizens-given-the-small-percentage-of-non-citizens-013}
ESS10_h=subset(ESS10_h, ESS10_h$ctzcntr %in% c(1))
print("ESS Citizenship")

ZA7649_h=subset(ZA7649_h, ZA7649_h$ctzcntr %in% c(1))
```

## Filter common countries
```{r filter-common-countries-014}
ESS10_h <- ESS10_h %>%
  filter(cntry %in% c("BE", "BG", "CZ", "EE", "FI", "FR",  "GR", "HR", "HU", "IE", "IT", "LT", "NL", "PT", "SI", "SK")) %>%
  mutate(
    Country_code = case_when(
      cntry == "BE" ~ 1,
      cntry == "BG" ~ 2,
      cntry == "CZ" ~ 3,
      cntry == "EE" ~ 4,
      cntry == "FI" ~ 5,
      cntry == "FR" ~ 6,
      cntry == "GR" ~ 7,
      cntry == "HR" ~ 8,
      cntry == "HU" ~ 9,
      cntry == "IE" ~ 10,
      cntry == "IT" ~ 11,
      cntry == "LT" ~ 12,
      cntry == "NL" ~ 13,
      cntry == "PT" ~ 14,
      cntry == "SI" ~ 15,
      cntry == "SK" ~ 16
    ),
    Country_code = factor(Country_code,levels = 1:16))

ZA7649_h <- ZA7649_h%>%
  filter(iso_h %in% c("BE", "BG", "CZ", "EE", "FI", "FR", "GR", "HR", "HU","IE", "IT", "LT", "NL", "PT", "SI", "SK")) %>%
  mutate(
    Country_code = case_when(
      iso_h == "BE" ~ 1,
      iso_h == "BG" ~ 2,
      iso_h == "CZ" ~ 3,
      iso_h == "EE" ~ 4,
      iso_h == "FI" ~ 5,
      iso_h == "FR" ~ 6,
      iso_h == "GR" ~ 7,
      iso_h == "HR" ~ 8,
      iso_h == "HU" ~ 9,
      iso_h == "IE" ~ 10,
      iso_h == "IT" ~ 11,
      iso_h == "LT" ~ 12,
      iso_h == "NL" ~ 13,
      iso_h == "PT" ~ 14,
      iso_h == "SI" ~ 15,
      iso_h == "SK" ~ 16
    ),
    Country_code = factor(Country_code,
       levels = 1:16))
```

### Trust
```{r trust-015}
# split missing values of ESS10$trstprl
ESS10_h=subset(ESS10_h, trstprl %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$trstprl
ESS10_h=ESS10_h %>%
  mutate(trstprl = case_when(
    trstprl %in% c(0,1,2,3,4,5) ~ 0,
    trstprl %in% c(6,7,8,9,10) ~ 1
  ))
ESS10_h$trstprl <- labelled(ESS10_h$trstprl,
                            c("No Trust" = 0,
                              "Trust" = 1))
# split missing values of ZA7649$qa6a_10
ZA7649_h=subset(ZA7649_h, qa6a_10 %in% c(1,2))

# Recoding values and labels of ZA7649$qa6a_10
ZA7649_h=ZA7649_h %>%
  mutate(qa6a_10 = case_when(
    qa6a_10 ==2 ~ 0,
    qa6a_10 ==1 ~ 1
  ))
ZA7649_h$qa6a_10 <- labelled(ZA7649_h$qa6a_10,
                             c("No Trust" = 0,
                               "Trust" = 1))

# split missing values of ESS10$trstlgl
ESS10_h=subset(ESS10_h, trstlgl %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$trstlgl
ESS10_h=ESS10_h %>%
  mutate(trstlgl = case_when(
    trstlgl %in% c(0,1,2,3,4,5) ~ 0,
    trstlgl %in% c(6,7,8,9,10) ~ 1
  ))
ESS10_h$trstlgl <- labelled(ESS10_h$trstlgl,
                            c("No Trust" = 0,
                              "Trust" = 1))


# split missing values of ZA7649$qa6a_3
ZA7649_h=subset(ZA7649_h, qa6a_3 %in% c(1,2))

# Recoding values and labels of ZA7649$qa6a_3
ZA7649_h=ZA7649_h %>%
  mutate(qa6a_3 = case_when(
    qa6a_3 ==2 ~ 0,
    qa6a_3 ==1 ~ 1
  ))
ZA7649_h$qa6a_3 <- labelled(ZA7649_h$qa6a_3,
                             c("No Trust" = 0,
                               "Trust" = 1))

# split missing values of ESS10$trstplc
ESS10_h=subset(ESS10_h, trstplc %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$trstplc
ESS10_h=ESS10_h %>%
  mutate(trstplc = case_when(
    trstplc %in% c(0,1,2,3,4,5) ~ 0,
    trstplc %in% c(6,7,8,9,10) ~ 1
  ))
ESS10_h$trstplc <- labelled(ESS10_h$trstplc,
                            c("No Trust" = 0,
                              "Trust" = 1))

# split missing values of ZA7649$qa6a_4
ZA7649_h=subset(ZA7649_h, qa6a_4 %in% c(1,2))

# Recoding values and labels of ZA7649$qa6a_4
ZA7649_h=ZA7649_h %>%
  mutate(qa6a_4 = case_when(
    qa6a_4 ==2 ~ 0,
    qa6a_4 ==1 ~ 1
  ))
ZA7649_h$qa6a_4 <- labelled(ZA7649_h$qa6a_4,
                            c("No Trust" = 0,
                              "Trust" = 1))

# split missing values of ESS10$trstprt
ESS10_h=subset(ESS10_h, trstprt %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$trstprt
ESS10_h=ESS10_h %>%
  mutate(trstprt = case_when(
    trstprt %in% c(0,1,2,3,4,5) ~ 0,
    trstprt %in% c(6,7,8,9,10) ~ 1
  ))
ESS10_h$trstprt <- labelled(ESS10_h$trstprt,
                            c("No Trust" = 0,
                              "Trust" = 1))

# split missing values of ZA7649$qa6a_2
ZA7649_h=subset(ZA7649_h, qa6a_2 %in% c(1,2))

# Recoding values and labels of ZA7649$qa6a_2
ZA7649_h=ZA7649_h %>%
  mutate(qa6a_2 = case_when(
    qa6a_2 ==2 ~ 0,
    qa6a_2 ==1 ~ 1
  ))
ZA7649_h$qa6a_2 <- labelled(ZA7649_h$qa6a_2,
                            c("No Trust" = 0,
                              "Trust" = 1))

# split missing values of ESS10$trstep
ESS10_h=subset(ESS10_h, trstep %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$trstep
ESS10_h=ESS10_h %>%
  mutate(trstep = case_when(
    trstep %in% c(0,1,2,3,4,5) ~ 0,
    trstep %in% c(6,7,8,9,10) ~ 1
  ))
ESS10_h$trstep <- labelled(ESS10_h$trstep,
                            c("No Trust" = 0,
                              "Trust" = 1))

# split missing values of ZA7649$qa6a_11
ZA7649_h=subset(ZA7649_h, qa6a_11 %in% c(1,2))

# Recoding values and labels of ZA7649$qa6a_11
ZA7649_h=ZA7649_h %>%
  mutate(qa6a_11 = case_when(
    qa6a_11 ==2 ~ 0,
    qa6a_11 ==1 ~ 1
  ))
ZA7649_h$qa6a_11 <- labelled(ZA7649_h$qa6a_11,
                            c("No Trust" = 0,
                              "Trust" = 1))

# split missing values of ESS10$trstun
ESS10_h=subset(ESS10_h, trstun %in% c(0,1,2,3,4,5,6,7,8,9,10))

# Recoding values and labels of ESS10$trstun
ESS10_h=ESS10_h %>%
  mutate(trstun = case_when(
    trstun %in% c(0,1,2,3,4,5) ~ 0,
    trstun %in% c(6,7,8,9,10) ~ 1
  ))
ESS10_h$trstun <- labelled(ESS10_h$trstun,
                           c("No Trust" = 0,
                             "Trust" = 1))


# split missing values of ZA7649$qa6a_12
ZA7649_h=subset(ZA7649_h, qa6a_12 %in% c(1,2))

# Recoding values and labels of ZA7649$qa6a_12
ZA7649_h=ZA7649_h %>%
  mutate(qa6a_12 = case_when(
    qa6a_12 ==2 ~ 0,
    qa6a_12 ==1 ~ 1
  ))
ZA7649_h$qa6a_12 <- labelled(ZA7649_h$qa6a_12,
                             c("No Trust" = 0,
                               "Trust" = 1))
```



# TARGET VARIABLES

## ESS latent traits
```{r ess-latent-traits-016}
# split missing values for items measuring latent traits
ESS10_h=subset(ESS10_h, imsmetn !=7 & imsmetn !=8 & imsmetn !=9)
ESS10_h=subset(ESS10_h, imdfetn !=7 & imdfetn !=8 & imdfetn !=9)
ESS10_h=subset(ESS10_h, impcntr !=7 & impcntr !=8 & impcntr !=9)

ESS10_h=subset(ESS10_h, imbgeco !=77 & imbgeco !=88 & imbgeco !=99)
ESS10_h=subset(ESS10_h, imueclt !=77 & imueclt !=88 & imueclt !=99)
ESS10_h=subset(ESS10_h, imwbcnt !=77 & imwbcnt !=88 & imwbcnt !=99)

# Recoding values and labels of ESS10_h$imbgeco
ESS10_h=ESS10_h %>%
  mutate(imbgeco = case_when(
    imbgeco %in% c(0,1) ~ 1,
    imbgeco %in% c(2,3,4) ~ 2,
    imbgeco %in% c(5,6,7) ~ 3,
    imbgeco %in% c(8,9,10) ~ 4
  ))

# Recoding values and labels of ESS10_h$imueclt
ESS10_h=ESS10_h %>%
  mutate(imueclt = case_when(
    imueclt %in% c(0,1) ~ 1,
    imueclt %in% c(2,3,4) ~ 2,
    imueclt %in% c(5,6,7) ~ 3,
    imueclt %in% c(8,9,10) ~ 4
  ))

# Recoding values and labels of ESS10_h$imwbcnt
ESS10_h=ESS10_h %>%
  mutate(imwbcnt = case_when(
    imwbcnt %in% c(0,1) ~ 1,
    imwbcnt %in% c(2,3,4) ~ 2,
    imwbcnt %in% c(5,6,7) ~ 3,
    imwbcnt %in% c(8,9,10) ~ 4
  ))

```

## Eurobarometer latent traits
```{r eurobarometer-latent-traits-017}
# split missing values for items measuring latent traits
ZA7649_h=subset(ZA7649_h, qb5_4 !=3 & qb5_4 !=4 & qb5_4 !=9)
ZA7649_h=subset(ZA7649_h, qb6_1 !=3 & qb6_1 !=4 & qb6_1 !=9)
ZA7649_h=subset(ZA7649_h, qb6_2 !=3 & qb6_2 !=4 & qb6_2 !=9)

ZA7649_h=subset(ZA7649_h, qb7_1 !=5  & qb7_1 !=9)
ZA7649_h=subset(ZA7649_h, qb7_2 !=5  & qb7_2 !=9)
ZA7649_h=subset(ZA7649_h, qb8_1 !=5  & qb8_1 !=9)
ZA7649_h=subset(ZA7649_h, qb8_2 !=5  & qb8_2 !=9)
```


# Subsetting by variables
## ESS
```{r ess-018}
ESS10_h_s <- ESS10_h %>%
  select(Country=cntry,
         Country_code=Country_code,
         Age=agea,
         Gender=gndr,
         Occupation=mnactic,
         Domicile=domicil,
         Subjective_income=hincfel,
         Political_Orientation=lrscale,
         Attachment_country=atchctr,
         Life_Satisfaction=stflife,
         Economy_Satisfaction=stfeco,
         Trust_Parliament=trstprl,
         Trust_Legalsystem=trstlgl,
         Trust_police=trstplc,
         Trust_politicalparties=trstprt,
         Trust_EU=trstep,
         Trust_UN=trstun,
         Allow_samerace=imsmetn,
         Allow_differentrace=imdfetn,
         Allow_poorcountries=impcntr,
         Impact_economy=imbgeco,
         Impact_cultural=imueclt,
         Impact_country=imwbcnt,
         Weights=anweight)%>%
  mutate(
    #   Numeriche
    Age = as.numeric(Age),
    Weights = as.numeric(Weights),

    #   Nominali (Categoriali)
    Country = as.factor(Country),
    Country_code=as.factor(Country_code),
    Gender = as.factor(Gender),
    Occupation = as.factor(Occupation),
    Political_Orientation = as.factor(Political_Orientation),
    Domicile = as.factor(Domicile),

    #   Ordinali (Specificando l'ordine)
    Subjective_income = factor(Subjective_income, levels = c(1, 2), ordered = TRUE),
    Attachment_country = factor(Attachment_country, levels = 1:4, ordered = TRUE),
    Life_Satisfaction = factor(Life_Satisfaction, levels = 1:4, ordered = TRUE),
    Economy_Satisfaction = factor(Economy_Satisfaction, levels = 1:4, ordered = TRUE),
    Trust_Parliament= factor(Trust_Parliament,levels = 0:1),
    Trust_Legalsystem=factor(Trust_Legalsystem,levels = 0:1),
    Trust_police=factor(Trust_police,levels = 0:1),
    Trust_politicalparties=factor(Trust_politicalparties,levels = 0:1),
    Trust_EU=factor(Trust_EU,levels = 0:1),
    Trust_UN=factor(Trust_UN,levels = 0:1),
    Allow_samerace=factor(Allow_samerace, levels = 1:4, ordered = TRUE),
    Allow_differentrace=factor(Allow_differentrace, levels = 1:4, ordered = TRUE),
    Allow_poorcountries=factor(Allow_poorcountries, levels = 1:4, ordered = TRUE),
    Impact_economy=factor(Impact_economy, levels = 1:4, ordered = TRUE),
    Impact_cultural=factor(Impact_cultural, levels = 1:4, ordered = TRUE),
    Impact_country=factor(Impact_country, levels = 1:4, ordered = TRUE),
  )

```

## Eurobarometer
```{r eurobarometer-019}
ZA7649_h_s <- ZA7649_h %>%
  select(Country=iso_h,
         Country_code=Country_code,
         Age=d11,
         Gender=d10,
         Occupation=d15a,
         Domicile=d25,
         Subjective_income=d60,
         Political_Orientation=d1,
         Attachment_country=qc1a_2,
         Life_Satisfaction=d70,
         Economy_Satisfaction=qa1a_2,
         Trust_Parliament=qa6a_10,
         Trust_Legalsystem=qa6a_3,
         Trust_police=qa6a_4,
         Trust_politicalparties=qa6a_2,
         Trust_EU=qa6a_11,
         Trust_UN=qa6a_12,
         Common_policy=qb5_4,
         Asylum_system=qb6_1,
         External_border=qb6_2,
         Imm_EU=qb7_1,
         Imm_NonEU=qb7_2,
         Im_contr=qb8_1,
         help_refugees=qb8_2,
         Weights=w1)%>%
  mutate(
    #   Numeriche
    Age = as.numeric(Age),
    Weights = as.numeric(Weights),

    #   Nominali (Categoriali)
    Country = as.factor(Country),
    Country_code=as.factor(Country_code),
    Gender = as.factor(Gender),
    Occupation = as.factor(Occupation),
    Political_Orientation = as.factor(Political_Orientation),
    Domicile = as.factor(Domicile),

    #   Ordinali (Specificando l'ordine)
    Subjective_income = factor(Subjective_income, levels = c(1, 2), ordered = TRUE),
    Attachment_country = factor(Attachment_country, levels = 1:4, ordered = TRUE),
    Life_Satisfaction = factor(Life_Satisfaction, levels = 1:4, ordered = TRUE),
    Economy_Satisfaction = factor(Economy_Satisfaction, levels = 1:4, ordered = TRUE),
    Trust_Parliament= factor(Trust_Parliament,levels = 0:1),
    Trust_Legalsystem=factor(Trust_Legalsystem,levels = 0:1),
    Trust_police=factor(Trust_police,levels = 0:1),
    Trust_politicalparties=factor(Trust_politicalparties,levels = 0:1),
    Trust_EU=factor(Trust_EU,levels = 0:1),
    Trust_UN=factor(Trust_UN,levels = 0:1),
    Common_policy=factor(Common_policy, levels = 1:2, ordered = TRUE),
    Asylum_system=factor(Asylum_system, levels = 1:2, ordered = TRUE),
    External_border=factor(External_border, levels = 1:2, ordered = TRUE),
    Imm_EU=factor(Imm_EU, levels = 1:4, ordered = TRUE),
    Imm_NonEU=factor(Imm_NonEU, levels = 1:4, ordered = TRUE),
    Im_contr=factor(Im_contr, levels = 1:4, ordered = TRUE),
    help_refugees=factor(help_refugees, levels = 1:4, ordered = TRUE),
    )
```

## COMBINED DATASET
```{r combined-dataset-020}
# Get common variables
common_vars <- intersect(names(ESS10_h_s), names(ZA7649_h_s))
# Specify the variables you want to impute
vars_ESS <- c("Allow_samerace", "Allow_differentrace",
         "Allow_poorcountries","Impact_economy",
         "Impact_cultural","Impact_country")
vars_ZA <- c("Common_policy", "Asylum_system",      "External_border","Imm_EU","Imm_NonEU","Im_contr","help_refugees")

# Add missing columns to ZA7649_h_s
for (v in vars_ESS) {
  ZA7649_h_s[[v]] <- NA
}

# Add missing columns to ESS10
for (v in vars_ZA) {
  ESS10_h_s[[v]] <- NA
}
# Combine the full list
all_vars <- union(common_vars, vars_ESS)
all_vars <- union(all_vars,vars_ZA)
# Subset the datasets
ess_data <- ESS10_h_s[, all_vars]
za_data  <- ZA7649_h_s[, all_vars]
ess_data$source <- "ESS"
za_data$source  <- "ZA"

# Combine into one dataset
combined_data <- rbind(ess_data, za_data)%>%
 mutate(
    Common_policy=factor(Common_policy, levels = 1:2, ordered = TRUE),
    Asylum_system=factor(Asylum_system, levels = 1:2, ordered = TRUE),
    External_border=factor(External_border, levels = 1:2, ordered = TRUE),
    Imm_EU=factor(Imm_EU, levels = 1:4, ordered = TRUE),
    Imm_NonEU=factor(Imm_NonEU, levels = 1:4, ordered = TRUE),
    Im_contr=factor(Im_contr, levels = 1:4, ordered = TRUE),
    help_refugees=factor(help_refugees, levels = 1:4, ordered = TRUE),
    )

rm(v,za_data,ess_data)

```

```{r remove-data-021}
rm(ESS10)
rm(ESS10_h)
rm(ZA7649)
rm(ZA7649_h)

```

## ########### LATENT TRAIT ESTIMATION ##########################

## LATENT COMMON TRAIT - TRUST - unidimensional IRT model

```{r latent-common-trait-trust-unidimensional-irt-model-022}
data_trust <- as.matrix(cbind(
  combined_data$Trust_Parliament,
  combined_data$Trust_Legalsystem,
  combined_data$Trust_police,
  combined_data$Trust_politicalparties,
  combined_data$Trust_EU,
  combined_data$Trust_UN
))
colnames(data_trust)=c("Trust_Parliament","Trust_Legalsystem","Trust_police","Trust_politicalparties","Trust_EU","Trust_UN")
```

#### Configural Invariant model
```{r configural-invariance-023}
mod_configural_trust=multipleGroup(data_trust,model = 1,group = combined_data$Country,itemtype = "graded",NCYCLES=1000, verbose=F)
```

```{r configural-invariance-024}
combined_data= cbind(combined_data,fscores(mod_configural_trust))
names(combined_data)[names(combined_data) == "F1"] <- "TRUST"
```


```{r configural-invariance-025}
par(mfrow = c(1, 2))
hist(combined_data$TRUST[combined_data$source=="ESS"],breaks = 20, main = "ESS TRUST", xlab = "trust", col = "lightblue")
hist(combined_data$TRUST[combined_data$source=="ZA"],breaks = 20, main = "EUROBAROMETER TRUST", xlab = "trust", col = "salmon")

rm(dataESS,dataESS_sorted,dataEUR,dataEUR_sorted)

```


# TARGET VARIABLES - bidimensiona IRT model
#### ALLOWING IMMIGRANTS: _Higher values indicate greater opposition towards immigrants_ - Immigrant Acceptance-Restriction Scale
* Allow many/few immigrants of same race/ethnic group as majority
* Allow many/few immigrants of different race/ethnic group from majority	* Allow many/few immigrants from poorer countries outside Europe

#### OPINION ON IMMIGRANTS: _Higher values indicate more positive opinion_ -Immigrant Contribution Perception
* Immigration bad or good for country's economy
* Country's cultural life undermined or enriched by immigrants
* Immigrants make country worse or better place to live

```{r opinion-on-immigrants-026}
data_AF <- as.matrix(cbind(
  combined_data$Allow_samerace,
  combined_data$Allow_differentrace,
  combined_data$Allow_poorcountries,
  combined_data$Impact_economy,
  combined_data$Impact_cultural,
  combined_data$Impact_country
))

data_AF<-data_AF[combined_data$source=="ESS",]
colnames(data_AF)=c("Allow_samerace","Allow_differentrace","Allow_poorcountries","Impact_economy","Impact_cultural","Impact_country")

modelAF <- '
Allow = 1-3
Feeling = 4-6
COV = Allow*Feeling
'
```

#### Configural Invariant model
```{r configural-invariance-027}
mod_configural_AF=multipleGroup(data_AF,model = modelAF,group = combined_data$Country[combined_data$source=="ESS"],itemtype = "graded",method='MHRM', verbose=T)
#summary(mod_configural_AF)
```


```{r configural-invariance-028}
combined_data= cbind(combined_data,rbind(fscores(mod_configural_AF),matrix(NA,dim(combined_data)[1]-dim(data_AF)[1],2)))
names(combined_data)[names(combined_data) == "Allow"] <- "ALLOW"
names(combined_data)[names(combined_data) == "Feeling"] <- "FEELING"
```

```{r configural-invariance-029}
par(mfrow = c(1, 2))

hist(combined_data$ALLOW[combined_data$source=="ESS"],breaks = 20, main = "ESS Allow scale", xlab = "allow", col = "red")
hist(combined_data$FEELING[combined_data$source=="ESS"],breaks = 20, main = "ESS Opinion scale", xlab = "opinion", col = "darkgreen")

```



## Eurobarometer latent traits - unidimensional IRT models

#### OPINION ON COMMON EU POLICIES ON IMMIGRATION (to correlate with imputed traits from ESS): _Higher values indicate higher acceptance of a common policy_
What is your opinion on each of the following statements? Please tell me for each statement, whether you are for it or against it.
* A common European policy on migration	EU PROPOSALS
* A common European Asylum system	EU PROPOSALS
* A reinforcement of EU external borders with more European border guards and coast guards	EU PROPOSALS

```{r opinion-on-common-eu-policies-on-immigration-030}
data_policy <- cbind(
  combined_data$Common_policy,
  combined_data$Asylum_system,
  combined_data$External_border
)
data_policy<-data_policy[combined_data$source=="ZA",]
colnames(data_policy) <- c("Common_policy", "Asylum_system", "External_border")
```

#### Configural Invariant model
```{r configural-invariance-031}
mod_configural_policy=multipleGroup(data_policy,model = 1,group = combined_data$Country[combined_data$source=="ZA"],itemtype = "graded",NCYCLES=1000, verbose=F)
```

```{r configural-invariance-032}
combined_data= cbind(combined_data,rbind(matrix(NA,dim(combined_data)[1]-dim(data_policy)[1],1),fscores(mod_configural_policy)))
names(combined_data)[names(combined_data) == "F1"] <- "COMMONPOLICY"

```


#### OPINION ON IMMIGRATION (control variable): lower values indicate more positive opinion
Please tell me whether each of the following statements evokes a positive or negative feeling for you?
* Immigration of people from other EU Member States
* Immigration of people from outside the EU

For each of the following statements, please tell me whether you totally agree, tend to agree, tend to disagree or totally disagree.
* Immigrants contribute a lot to (OUR COUNTRY)
* (OUR COUNTRY) should help refugees

```{r opinion-on-immigration-033}
data_feeling <- cbind(
  combined_data$Imm_EU,
  combined_data$Imm_NonEU,
  combined_data$Im_contr,
  combined_data$help_refugees
)
data_feeling<-data_feeling[combined_data$source=="ZA",]
colnames(data_feeling) <- c("Imm_EU", "Imm_NonEU", "Im_contr","help_refugees")
```

#### Configural Invariant model
```{r configural-invariance-034}
mod_configural_feeling=multipleGroup(data_feeling,model = 1,group = combined_data$Country[combined_data$source=="ZA"],itemtype = "graded",NCYCLES=1000, verbose=F)
```

```{r configural-invariance-035}
#summary(mod_configural_feeling)
combined_data= cbind(combined_data,rbind(matrix(NA,dim(combined_data)[1]-dim(data_feeling)[1],1),fscores(mod_configural_feeling)))
names(combined_data)[names(combined_data) == "F1"] <- "FEELING_ZA"

```

```{r configural-invariance-036}
par(mfrow = c(1, 2))

hist(combined_data$COMMONPOLICY[combined_data$source=="ZA"],breaks = 20, main = "Eurobarometer: Policy scale", xlab = "common policy", col = "red")
hist(combined_data$FEELING_ZA[combined_data$source=="ZA"],breaks = 30, main = "Eurobarometer: Opinion scale", xlab = "trust", col = "darkgreen")


```


# Clean Workspace
```{r clean-workspace-037}
rm(data_AF, data_trust, data_policy, data_feeling,ESS10_h_s,ZA7649_h_s, all_vars,common_vars,modelAF,vars_ESS,vars_ZA)
```

```{r clean-workspace-038}
save.image("ENCLOSE_harmonized_data.RData")
```
